Geometric variational problems

### 2020-11-05 NSF2020/Description/specificaim3_ver2.tex
working on aim 3

\subsection{Specific Aim 3: Efficient high-order numerical methods for
large-scale simulations}
\label{subsec:specific_aim_3}
% ----------------------------------------------------------------------
The HAP model requires solving the exterior Dirichlet problem of the
screened Laplace equation and the mobility problem of the Stokes flow at
each time step. If discretizing these equations with stencil-based
numerical methods, the computational domain must be truncated, the
volume must be discretized, and artificial boundary conditions must be
imposed. The artificial boundary conditions introduces additional error,
and discretizing the volume results in an excessively large linear
system. Moreover, it is difficult to obtain high-order discretizations
when the boundary surfaces are irregular. Since we are solving
constant-coefficient elliptic PDEs, we represent the solution of the
PDE of interest as a layer potential 
\begin{align}
  \label{eq:LP}
  u(x) = \int_{\partial\Omega} K(x,y) \sigma(y) ds_y,
\end{align}
where $\sigma$ is an unknown density function, and $K(x,y)$ is formed
with derivatives of the fundamental solution of the underlying PDE.  By
matching the boundary condition at $x_0 \in \partial\Omega$ with the
limit of equation~\eqref{eq:LP} as $x\rightarrow x_0 \in
\partial\Omega$, a BIE for the density function is formed.
Numerical methods to solve BIE has several advantages over their
PDE-based counterpart: by construction, $u$ satisfies far-field
conditions; only $\partial\Omega$ needs to be discretized; carefully
chosen kernels $K(x,y)$ result in a well-conditioned linear system that
requires a mesh-independent number of GMRES iterations to converge; and
high-order or spectral accuracy is attainable by using appropriate
quadrature methods. In our previous work~\cite{Fu2018_SIAM} we
demonstrate that BIEs are a powerful tool to simulate two-dimensional
suspensions of amphiphilic particles. In addition to improving
two-dimensional simulations, we will extend the results to three
dimensions using a standard three-dimensional BIE of the screened
Laplace equation~\cite{ying_2006} and a well-conditioned BIE of the
mobility problem~\cite{manasthesis}.

The most significant challenges of a BIE formulation include solving
dense linear systems, developing preconditioning strategies, and
developing quadrature methods for integrands that are nearly singular.
These challenges are described in Sections~\ref{subsec:NumericalIssues}.

% ----------------------------------------------------------------------
\subsubsection{Numerical issues}
\label{subsec:NumericalIssues}
\todo[inline]{Preconditioning}
\todo[inline]{Adaptive high-order time stepping}

Discretizations of carefully chosen BIEs can be solved with a
mesh-independent number of GMRES
iterations~\cite{cam-ips-kel-mey-xue1996}. Therefore, the required CPU
time is proportional to the cost of a matrix-vector multiplication that
can be done in optimal or near-optimal time with the fast multiple
method (FMM)~\cite{fmm5} and its extensions~\cite{fmm1, fmm2, fmm3,
fmm4, fmm6, fmm7, fmm8}. Alternatively, fast direct solvers can be used
to avoid any iteration~\cite{fds1, fds2, fds3, fds4, fds5, fds6, fds7,
fds8, ho2016cpam2, ho2016cpam1, minden2016, minden2017siammms}, but
direct solvers often require a large amount of initial overhead, and
this makes them less useful for problems with moving geometries. Alternatively, techniques in fast direct solvers can be used to develop
efficient preconditioning strategies.

@@

% ----------------------------------------------------------------------
\subsubsection{High-order adaptive time stepping}
\label{subsec:timeStepping}
\todo[inline]{Put contact in here as well}

Even with accurate quadrature and time stepping methods (see
Section~\ref{subsec:NumericalIssues}) methods, rigid body simulations
can still result in unphysical contact. In our previous work a
Leonard-Jones potential is used to keep the particles from
overlapping~\cite{Fu2018_SIAM}. Such steep steric interaction at short
ranges introduces great numerical stiffness that limit the time step
size.

@@

% ----------------------------------------------------------------------
\subsubsection{Fluctuating hydrodynamics}
Using DLP formulation for fluctuating hydrodynamics so that we have a
second-kind BIE

>>>>

\subsection{Specific Aim 3: Efficient high-order numerical methods for
large-scale simulations}
\label{subsec:specific_aim_3}
% ----------------------------------------------------------------------
The HAP model requires solving the exterior Dirichlet problem of the
screened Laplace equation and the mobility problem of the Stokes flow at
each time step in the complex domains such as Figure~\ref{fig:domain}.
If discretizing these equations with stencil-based numerical methods,
the computational domain must be truncated, the volume must be
discretized, and artificial boundary conditions must be imposed. The
artificial boundary conditions introduces additional error, and
discretizing the volume results in an excessively large linear system.
Moreover, it is difficult to obtain high-order discretizations when the
boundary surfaces are irregular. Since we are solving
constant-coefficient elliptic PDEs, we represent the solution of the PDE
of interest as a layer potential 
\begin{align}
  \label{eq:LP}
  u(x) = \int_{\partial\Omega} K(x,y) \sigma(y) ds_y,
\end{align}
where $\sigma$ is an unknown density function, and $K(x,y)$ is formed
with derivatives of the fundamental solution of the underlying PDE. By
matching the boundary condition at $x_0 \in \partial\Omega$ with the
limit of equation~\eqref{eq:LP} as $x\rightarrow x_0 \in
\partial\Omega$, a BIE for the density function is formed.
Numerical methods to solve BIE has several advantages over their
PDE-based counterpart: by construction, the layer
potential~\eqref{eq:LP} satisfies far-field conditions; only
$\partial\Omega$ needs to be discretized; carefully chosen kernels
$K(x,y)$ result in a well-conditioned linear system that requires a
mesh-independent number of GMRES iterations to solve; and high-order or
spectral accuracy is attainable by using appropriate quadrature methods.
In our previous work~\cite{Fu2018_SIAM} we demonstrate that BIEs are a
powerful tool to simulate two-dimensional suspensions of amphiphilic
particles. In addition to improving two-dimensional simulations, we will
extend the results to three dimensions using a standard
three-dimensional BIE of the screened Laplace equation~\cite{ying_2006}
and a well-conditioned BIE of the mobility problem~\cite{manasthesis}.

The most significant challenges of a BIE formulation include solving
dense linear systems, developing preconditioning strategies, and
developing quadrature methods for integrands that are nearly singular.
These challenges are described in section~\ref{subsec:NumericalIssues}.
Then, in section~\ref{subsec:timeStepping}, time stepping strategies are
described, and stable BIE formulations for fluctuating hydrodynamics are
in section~\label{subsec:fluctuating}.

% ----------------------------------------------------------------------
\subsubsection{Numerical issues}
\label{subsec:NumericalIssues}

Discretizations of carefully chosen BIEs can be solved with a
mesh-independent number of GMRES
iterations~\cite{cam-ips-kel-mey-xue1996}. Therefore, the required CPU
time is proportional to the cost of a matrix-vector multiplication that
can be done in optimal or near-optimal time with the fast multiple
method (FMM)~\cite{fmm5} and its extensions~\cite{fmm1, fmm2, fmm3,
fmm4, fmm6, fmm7, fmm8}. PI Quaife is experienced with applying FMMs for
both the Stokes equation~\cite{qua-bir2014, bys-sha-qua2020} and the
screened Laplace equation~\cite{kro-qua2011, qua2011}. Alternatively,
fast direct solvers can be used to avoid any iteration~\cite{fds1, fds2,
fds3, fds4, fds5, fds6, fds7, fds8, ho2016cpam2, ho2016cpam1,
minden2016, minden2017siammms}, but direct solvers often require a large
amount of initial overhead, and this makes them less useful for problems
with moving geometries.  Alternatively, techniques in fast direct
solvers can be used to develop efficient preconditioning strategies. PI
Quaife used the inverse Fast Multipole Method
(IFMM)~\cite{cou-pou-dar2017, qua-cou-dar2018} to precondition Stokes
equations in a porous media, and a suite of preconditioners ranging from
block-diagonal to the IFMM will be considered.

@@

% ----------------------------------------------------------------------
\subsubsection{Avoiding contact with adaptive time stepping and
repulsion}
\label{subsec:timeStepping}
\todo[inline]{Put contact in here as well}
\todo[inline]{Adaptive high-order time stepping}

Even with accurate quadrature and time stepping methods, rigid body
simulations can still result in unphysical contact. We proposed to use
two techniques to avoid unphysical contact: high-order adaptive time
stepping and a repulsion force. PI Quaife developed a high-order
adaptive time stepping method for hydrodynamic
suspensions~\cite{qua-bir2016} and he has used it with colleagues to
investigate mixing and adhesion in suspensions~\cite{qua-vee-you2019,
kab-qua-bir2017}.

In our previous work a
Leonard-Jones potential is used to keep the particles from
overlapping~\cite{Fu2018_SIAM}. Such steep steric interaction at short
ranges introduces great numerical stiffness that limit the time step
size. 
\cite{bys-sha-qua2020}

@@

% ----------------------------------------------------------------------
\subsubsection{Fluctuating hydrodynamics}
\label{subsec:fluctuating}
Using DLP formulation for fluctuating hydrodynamics so that we have a
second-kind BIE
